Flabby and injective objects in toposes

### 2022-02-16 paper-flabby-objects.tex
Finished section on enough injectives

\begin{prop}
\begin{enumerate}
\item Assuming the law of excluded middle, divisible abelian groups satisfy
the Baer criterion.
\item Assuming Zorn's lemma, abelian groups which satisfy the Baer criterion
are injective.
\end{enumerate}
\end{prop}

@@

Assuming Zorn's lemma, let~$I$ be an abelian group satisfying the Baer
criterion, let~$i : A \to B$ be a linear injection and let~$f : A \to I$ be a
linear map. The poset of partial extensions contains suprema of chains and
hence a maximal partial extension~$f_0 : A_0 \to I$ with~$A \subseteq A_0
\subseteq B$. To verify that~$A_0 = B$, let an element~$x \in B$ be given.
By the Baer condition, the linear map~$g : (A_0 : x) \to I$ defined on the
ideal~$(A_0 : x) = \{ n \in \ZZ \,|\, nx \in A_0 \} \subseteq \ZZ$ given by~$g(n) =
f_0(nx)$ can be extended to a linear map~$\overline{g} : \ZZ \to I$.
Hence~$f_0$ can be extended to the map~$A_0 + (x) \to I$ given by~$u + nx
\mapsto f_0(u) + \overline{g}(n)$. By maximality~$A_0 + (x) = A_0$, hence~$x \in A_0$.
\end{proof}

\begin{prop}
\begin{enumerate}
\item Any abelian group embeds into a divisible abelian group.
\item Any abelian group embeds into an abelian group satisfying the Baer
condition.
\end{enumerate}
\end{prop}

@@

The second part is more involved and requires, similar to the proof of
Theorem~\ref{thm:enough-flabby-modules}, a strengthening of the Baer condition
where the required extensions are explicitly given. As a first step, let~$A$ be
an abelian group which is a sheaf for a modal operator~$\nabla$ such
that~$\nabla(\varphi \vee (\varphi \Rightarrow \nabla\bot))$. Let~$i : \aaa
\hookrightarrow \ZZ$ be the inclusion of an ideal of~$\ZZ$ and let~$f : \aaa
\to A$ be a linear map. Applying the sheafification functor~$(\cdot)^{++}$, we
have the following commutative diagram:
\[ \xymatrix{
  \aaa \ar@{^{(}->}[r]^i\ar[d]\ar@/_2pc/[dd]_f & \ZZ \ar[d] \\
  \aaa^{++} \ar@{^{(}->}[r]^{i^{++}}\ar@{-->}[d] & \ZZ^{++}\ar@{-->}@/^/[ld]^{f'} \\
  A
} \]
The given map~$f$ factors (uniquely) over the map~$\aaa \to \aaa^{++}$ since~$A$ is a
sheaf. The map~$i^{++}$ is still injective as sheafification is exact. The two
top vertical maps need not be injective, as~$\ZZ$ need not be separated
for~$\nabla$; but this issue XXX
\end{proof}

\begin{prop}Assuming Zorn's lemma:
\begin{enumerate}
\item Any abelian group embeds into an injective abelian group.
\item Any module embeds into an injective module.
\item XXX Grothendieck topos?
\end{enumerate}
\end{prop}

>>>>

\begin{prop}
\label{prop:embed1}
\begin{enumerate}
\item Assuming the law of excluded middle, all divisible abelian groups satisfy
the Baer criterion.
\item Assuming Zorn's lemma, all abelian groups which satisfy the Baer criterion
are injective.
\end{enumerate}
\end{prop}

@@

Assuming Zorn's lemma, let~$I$ be an abelian group satisfying the Baer
criterion, let~$i : A \to B$ be a linear injection and let~$f : A \to I$ be a
linear map. The poset of partial extensions contains suprema of chains and
hence a maximal partial extension~$f_0 : A_0 \to I$ with~$A \subseteq A_0
\subseteq B$.

To verify that~$A_0 = B$, let an element~$x \in B$ be given.
By the Baer condition, the linear map~$g : (A_0 : x) \to I$ defined on the
ideal~$(A_0 : x) = \{ n \in \ZZ \,|\, nx \in A_0 \} \subseteq \ZZ$ given by~$g(n) =
f_0(nx)$ can be extended to a linear map~$\overline{g} : \ZZ \to I$.
Hence~$f_0$ can be extended to the map~$A_0 + (x) \to I$ given by~$u + nx
\mapsto f_0(u) + \overline{g}(n)$. By maximality~$A_0 + (x) = A_0$, hence~$x \in A_0$.
\end{proof}

\begin{lemma}\label{lemma:q-mod-z}
Let~$\nabla$ be a modal operator such
that~$\nabla(\varphi \vee (\varphi \Rightarrow \nabla\bot))$ and
let~$(\cdot)^+$ denote the plus construction with respect to~$\nabla$. Then:
\begin{enumerate}
\item The sheafification~$(\QQ/\ZZ)^{++}$ satisfies the Baer condition.
\item Assuming Zorn's lemma, given an element~$x$ of an abelian group~$A$,
there is a linear map~$g : A \to (\QQ/\ZZ)^{++}$ such that~$g(x) = 0
\Rightarrow \nabla(x=0)$.
\end{enumerate}
\end{lemma}

\begin{proof}Let~$i : \aaa \hookrightarrow \ZZ$ be the inclusion of an ideal
and let~$f : \aaa \to A$ be a linear map as in the solid part of the following
commutative diagram:
\[ \xymatrix{
  \aaa \ar@{^{(}->}[r]^i\ar[d]\ar@/_2pc/[dd]_f & \ZZ \ar[d] \\
  \aaa^{++} \ar@{^{(}->}[r]^{i^{++}}\ar@{-->}[d] & \ZZ^{++}\ar@{-->}@/^/[ld]^{f'} \\
  (\QQ/\ZZ)^{++}
} \]
The given map~$f$ factors (uniquely) over the map~$\aaa \to \aaa^{++}$
since~$(\QQ/\ZZ)^{++}$ is a sheaf. The map~$i^{++}$ is still injective as
sheafification is exact. (The two top vertical maps need not be injective,
as~$\ZZ$ need not be separated for~$\nabla$.) Hence we are reduced to an
extension problem in the subtopos of~$\nabla$-sheaves, which is solvable
because this subtopos is Boolean (Proposition~\ref{prop:embed1}(1)). More
explicitly, the extension~$f'$ maps~$1 \in \ZZ$ to the gluing of
\begin{multline*}
  \{ 0 \,|\, \forall x \in \aaa. \nabla(x = 0) \} \cup {}\\
  \{ \tfrac{1}{d} v \,|\, d \geq 1, v \in [0,1),
    f(d) = \underline{v}, \forall x \in \aaa. \nabla(x \in \aaa)
    \Leftrightarrow \nabla(\exists n \in \ZZ. x = nd) \}.
\end{multline*}

For the second part, let~$A$ be an abelian group and let~$x \in A$ be an
element. Let~$i : \ZZ/(0:x) \to A$ be the linear map with~$[1] \mapsto x$,
where~$(0:x)$ is the ideal~$\{ n \in \ZZ \,|\, nx = 0 \}$. In the Boolean
subtopos of~$\nabla$-sheaves, this ideal is either the zero ideal or the
principal ideal~$(d)$ of some positive generator. Let~$h : \ZZ/(0:x) \to
(\QQ/\ZZ)^{++}$ be the linear map sending~$[1]$ to the respective gluing such
that in the first case~$h([1]) = [\tfrac{1}{2}]$ and in the second case~$h([1])
= [\frac{1}{d}]$. Assuming Zorn's lemma such that
Proposition~\ref{prop:embed1}(2) applies, there is an extension~$\overline{h} :
A \to (\QQ/\ZZ)^{++}$. This map is the desired map. If~$\overline{h}(x) =
h([1]) = 0$, then in the subtopos~$d = 1$, hence~$\nabla(x = 0)$.
\end{proof}

\begin{prop}\label{prop:embed2}
\begin{enumerate}
\item Every abelian group embeds canonically into a divisible abelian group.
\item Every abelian group maps canonically to an abelian group functionally satisfying the Baer
condition. Assuming Zorn's lemma, this map is an embedding.
\end{enumerate}
\end{prop}

@@

The second part is more involved and requires, similar to the proof of
Theorem~\ref{thm:enough-flabby-modules}, a strengthening of the Baer condition
where the required extensions are explicitly given. As a first step, let~$A$ be
an abelian group which is separated for a modal operator~$\nabla$ such
that~$\nabla(\varphi \vee (\varphi \Rightarrow \nabla\bot))$. The codomain of
the canonical map
\[ A \longrightarrow \prod_{g : A \to (\QQ/\ZZ)^{++}} (\QQ/\ZZ)^{++},\
  x \longmapsto (g(x))_g \]
is a product of abelian groups which functionally satisfy the Baer criterion by
Lemma~\ref{lemma:q-mod-z}(1) and hence satisfies the Baer condition itself.
Assuming Zorn's lemma, this map is injective by Lemma~\ref{lemma:q-mod-z}(2)
and~$\nabla$-separatedness of~$A$.

For the general case, let an abelian group~$A$ be given. Denoting
by~$(\cdot)^{+_x}$ the plus construction with respect to the modal
operator~$\nabla_x$ with~$\nabla_x\varphi \defeqv ((\varphi \Rightarrow x
= 0) \Rightarrow x = 0)$, the codomain of the canonical map
\[ A \longrightarrow \prod_{x \in A} \prod_{g : A \to (\QQ/\ZZ)^{+_x+_x}}
(\QQ/\ZZ)^{+_x+_x} \]
satisfies the Baer criterion and is injective as in the proof of
Theorem~\ref{thm:enough-flabby-modules}.
\end{proof}

\begin{prop}Assuming Zorn's lemma:
\begin{enumerate}
\item Every abelian group embeds canonically into an injective abelian group.
\item Every module embeds canonically into an injective module.
\item Every module in any Grothendieck topos embeds canonically into an injective module.
\end{enumerate}
\end{prop}

\begin{proof}The first claim is by Proposition~\ref{prop:embed2}.

The second claim follows purely formally as in the proof of
Proposition~\ref{prop:basics-injective}(3), using the adjunction between the
category of modules and the category of abelian groups.

For the third claim, let~$\E$ be a Grothendieck topos and let~$\pi : \F \to \E$ be its Barr
covering \XXX{ref}. Zorn's lemma passes from the metatheory to~$\F$ since~$\F$ is a
localic topos. Hence the first part applies in~$\F$. Purely formally, using the
adjunction between~$\mathrm{Ab}(\F)$ and~$\mathrm{Ab}(\E)$, there are enough
injective abelian groups in~$\E$. The existence of enough injective modules
in~$\E$ follows from internalizing the second claim in~$\E$.
\end{proof}

\XXX{cite whom for the general argument?}
